---
title: "titanic"
output: titanic-R
---

#Package
```{r}
# Load packages
library('ggplot2') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
# library('randomForest') # classification algorithm
library('data.table')#dt
library('plotly')
library('highcharter')
require(corrplot)
```

##plotly margin

```{r}
# Margin for btm padding for plotly
m <- list(
  b = 150,
  pad = 4
)
```


#Read Data
```{r}
train <- data.table(read.csv('../input/train.csv', stringsAsFactors = F, na.strings=""))
test  <- data.table(read.csv('../input/test.csv', stringsAsFactors = F, na.strings=""))

full  <- bind_rows(train, test) # bind training & test data

full<- data.table(rbind(train[,Identity:="train"], test[,Identity:="test"], fill = T))
full$Identity<- as.factor(full$Identity)

```



# Glimpse, Summary, Column Name
Making column name more sensible and intuitive
```{r}
# setnames(full, c("Pclass", "SibSp", "Parch"),c("PassengerClass", "N_SibSpouse", "N_ParentsChild"))

glimpse(full)

summary(full)

# names_factors=c('PassengerClass', 'Sex', 'Embarked')
# for (col in names_factors) set(full, j=col, value=as.factor(full[[col]])) #changing selected column types
# http://stackoverflow.com/questions/7813578/convert-column-classes-in-data-table?rq=1#comment31200110_20808945

## Other methods to change column types in data table
# full <- full[, PassengerClass:=as.factor(PassengerClass), ']
# dtnew <- dt[, lapply(.SD, as.character), by=ID]

## Method to identify NA if it is not specified in read.csv
# full[,Embarked==""] <- NA
```



#Missing values



```{r}
# Exploring NA values
fullNA.dt<-full %>%
  .[,list(
    # Numbers = .N,
          "% Pclass" = sum(!is.na(Pclass))/.N,
          "% Name" = sum(!is.na(Name))/.N,
          "% Sex" = sum(!is.na(Sex))/.N,
          "% Age" = sum(!is.na(Age))/.N,
          "% SibSp" = sum(!is.na(SibSp))/.N,
          "% Parch" = sum(!is.na(Parch))/.N,
          "% Ticket" = sum(!is.na(Ticket))/.N,
          "% Fare" = sum(!is.na(Fare))/.N,
          "% Cabin" = sum(!is.na(Cabin))/.N,
          "% Embarked" = sum(!is.na(Embarked))/.N
          ),
    by=list(Survived)]



fullNA_melt.dt<- fullNA.dt %>%
  melt(., id = 1) 

# plot
fullNA_melt.dt$Survived[is.na(fullNA_melt.dt$Survived)]  <- 'U'
fullNA_melt.dt <- fullNA_melt.dt[, Survived:=as.character(Survived)]


fullNA_melt.dt%>%
  plot_ly(.,
          x = ~variable,
          y = ~value,
          color = ~Survived,
          name = "Survived",
          type = "bar"
  ) %>%
  layout(margin = m)


```

# Title
Other kernel such as Megan Risdal - Exploring on the Titanic does a good job of explaining this
https://www.kaggle.com/mrisdal/exploring-survival-on-the-titanic
Basically, the name column contained title that is associated with social class and thereby survival rate
```{r}
# Grab title from passenger names
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)

# Show title counts by sex
table(full$Sex, full$Title)


# Titles with very low cell counts to be combined to "rare" level
rare_title <- c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')

# Also reassign mlle, ms, and mme accordingly
full$Title[full$Title == 'Mlle']        <- 'Miss' 
full$Title[full$Title == 'Ms']          <- 'Miss'
full$Title[full$Title == 'Mme']         <- 'Mrs' 
full$Title[full$Title %in% rare_title]  <- 'Rare Title'

# Show title counts by sex again
table(full$Sex, full$Title)

# Plotting to visuallise the effect on suvival based on the different in title
full %>%
  plot_ly(x = ~Title,
          color= ~as.factor(Survived),
             type = "histogram",
             histnorm = "probability")

#note that survived ~ 2 is NA

full$Title<- as.factor(full$Title)
```


```{r}
# Show title counts by sex
# table(train.dt$Title,train.dt$Survived)

Train_Title_Survival.dt<-full[!is.na(Survived)] %>%
  .[,list(
     Population = .N,
     SurviveP = sum((Survived))/.N
          ),
    by=list(Title)]%>%
  # Order the Survival by their mean value
  .[order(-SurviveP)]



Train_Title_Survival.dt$Title <- factor(Train_Title_Survival.dt$Title, 
                                 levels = Train_Title_Survival_order.dt[["Title"]])


Train_Title_Survival.dt%>%
  plot_ly(.,
          x = ~Title,
          y = ~SurviveP,
          color = ~Title,
          name = "Survived",
          type = "bar"
  ) %>%
  layout(margin = m)

```

# N_Sib&Spouse & N_Parents&Child
These column's relevence will be explore during fitting with bootstrap

```{r}
full %>%
  plot_ly(x = ~SibSp,
          color= ~as.factor(Survived),
             type = "histogram",
             histnorm = "probability")

full %>%
  plot_ly(x = ~Parch,
          color= ~as.factor(Survived),
             type = "histogram",
             histnorm = "probability")

#note that survived ~ 2 is NA


```
It seems it is helpful to have a fellow family on board while the titanic sink, as those with 2, 3 family members tend to survive. Though, it may be better if these columns are combined as **N_family**


## N_family 
```{r}
# Create a family size variable including the passenger themselves
full$Fsize <- full$SibSp + full$Parch + 1

full_Fsize_Survival.dt<-full[!is.na(Survived)]%>%
  .[,list(
     Population = .N,
     SurviveP = sum(Survived)/.N
          ),
    by=list(Fsize)]%>%
  .[order(Fsize)]

full_Fsize_Survival.dt%>%
  plot_ly(.,
          x = ~Fsize,
          y = ~SurviveP,
          color = ~Population,
          name = "Survived",
          type = "bar"
  ) %>%
  layout(margin = m)
```


# Age
There is significant amount of NA in Age column but like sex and titles, it should play a role in determining a person's survival rate.
```{r}
# Construct the AgeGroup
breaks <- c(0,5,18,25,35,45,65,Inf)
tempt <- full[,list(SurvivedR=mean(as.numeric(Survived), na.rm=TRUE)),
     by=list(AgeGroup=cut(Age,breaks=breaks))][order(AgeGroup)]
# Adapted from 
# https://stackoverflow.com/questions/27107454/r-data-table-group-data-by-custom-range-for-example-18-18-25-65


# Plotting the result
tempt %>%
  plot_ly(x = ~AgeGroup,
          y = ~SurvivedR,
             type = "bar")

# Create a column for AgeGroup
full[,AgeGroup:=cut(Age,breaks=breaks)]


# It seems children are probably prioritized for the rescue effort, while the older cohorts fared worse during titanic
# maybe using kneighbors to setup AgeGroup might be a better idea

```

## Correlation
```{r}
# for correlations to work, the columns should either numerical or categorical(factors)

character_columns<-c("Embarked", "Sex", "Title")
fct_records
full[, (character_columns):=lapply(.SD, as.factor), .SDcols= character_columns]


full[Identity=="train",]%>%
  .[, (character_columns):=lapply(.SD, function(x) as.numeric(as.factor(x))), .SDcols= character_columns]%>%
  #seleciting columns for correlation
  .[, .(Survived, Pclass, Sex, Age, SibSp, Parch, Embarked, Title)]%>%
  cor(use="complete.obs")%>%
  corrplot(type ="lower", diag=FALSE)

install.packages("corrplot")
```

